---
title: "Visualize COVID-hospitalization filter, with chart-reviewed patients"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Written by Jeffrey Klann, PhD
Last Updated 4/20/21

This notebook will calculate the number of COVID hospitalizations by day and week when excluding patients who had a lab order for LDH, PaO2, PaCO2, or procalcitonin +/- 2 days around hospitalization, which is hypothesized (and shown on a small labeled dataset) to filter out most patients that have a hospitalization unrelated to COVID. This is only an exploratory analysis and this is not intended to be the final filter. That will come in time with chart review and machine learning.

Now it includes the ability to visualize chart-reviewed patients on the graph, loaded in an optional file called LocalPatientSeverityIcu with two columns: a patient_num and an admitted_for_covid flag (True, False, or NA)

Also it is now very easy to configure the filter used by altering fource_filter below.

To run this:
1) Install the plyr package if needed - install.packages("plyr")
2) Create the LocalPatientSeverityIcu.csv file if you would like to plot chart-reviewed points - see the comments near line 70
3) Set your threshold and blur below. Site ID is gleaned from the R package.
4) Change your filter below if you would like.
5) Run the code.

Future improvement list:
* Blur/remove summary stats count
* Do dots on percentage graph?


```{r}
library(tidyverse)
library(lubridate)
library(grid)
library(glue)
library(directlabels)

# Get params from FourCePhase2.1Data
# data_dir <- "../4ce/Input"
# currSiteId <- "upenn"
# fake_chart_review <- TRUE
# threshold <- 10
# blur <- 0

data_dir <- FourCePhase2.1Data::getInputDataDirectoryName()
currSiteId <- FourCePhase2.1Data::getSiteId()
  
# 8/13/21 - jgk - set directory for 2.2 version if desired
#data_dir <- '/4ceData_x2/4ceData_x2/Local/Input'

fake_chart_review <- FALSE
# SITES: SET THESE VALUES IF YOU PLAN TO SHARE YOUR RESULTS
threshold <- 3
blur <- 3

# Also configurable: what filter will we use to do our analysis?
# LDH, PaO2, PaCO2, or procalcitonin
fource_filter <- c("33959-8", "2532-0", "2019-8", "2703-7")
# fource_filter <- c('48065-7') # ddimer FEU

```

```{r}
CurrSiteId <- toupper(currSiteId)
# Load the data
clin_raw <-
  readr::read_csv(
    file.path(data_dir, "LocalPatientClinicalCourse.csv"),
    col_types = list(patient_num = readr::col_character())
  )
obs_raw <-
  readr::read_csv(
    file.path(data_dir, "LocalPatientObservations.csv"),
    col_types = list(patient_num = readr::col_character())
  )
demo_raw <-
  readr::read_csv(
    file.path(data_dir, "LocalPatientSummary.csv"),
    col_types = list(patient_num = readr::col_character()),
    na = c("1900-01-01", "1/1/1900")
  ) %>%
  # Fix for datetimes (Oracle code?)
  mutate(admission_date = as_date(admission_date))

# Optional - 
# Load table with chart review flag data
# This script must have at least three columns: 
#    patient_num : the patient_num from the 2.1 study cohort, not necessarily i2b2
#    admitted_for_covid : chart-reviewed admission flag, 1 if chart-reviewed as admitted for COVID, 0 if chart-reviewed as not admitted for COVID, and 2 for maybe.
#    severe_covid : chart-reviewed severity flag, 1 if chart-reviewed severe illness, 0 if chart-reviewed non-severe illness, NULL otherwise

if (fake_chart_review) { # fake number
  icu_raw <- tibble(
    patient_num = as.character(1:50),
    admitted_for_covid = c(rep(TRUE, 40), rep(FALSE, 10))
  )
} else {
  icu_raw <- tibble(patient_num = "x", admitted_for_covid = NA)
  if (file.exists(file.path(data_dir, "LocalPatientAdmissionChartReview.csv"))) {
    icu_raw <-
      readr::read_csv(
        file.path(data_dir, "LocalPatientAdmissionChartReview.csv.csv"),
        col_types = list(patient_num = readr::col_character())
      )
  }
}
```

This part calculates stuff. If you've loaded icu_raw, calculate the number of patients with chart-reviewed COVID-admission or chart-reviewed non-COVID-admission

```{r}
round_any <- function(x, accuracy, f = round) {
  f(x / accuracy) * accuracy
}

# Blurring function from Trang. Thanks Trang!
blur_it <- function(df, vars, blur_abs, mask_thres) {
  # Obfuscate count values.
  # If blurring range is +/-3, or blur_abs = 3,
  # the count receive a small addition of a random number from -3 to 3.
  # If a count is less than mask_thres, set that count to 0.

  for (var in vars) {
    var <- sym(var)
    blur_vec <- sample(seq(-blur_abs, blur_abs), nrow(df), replace = TRUE)
    df <- df %>%
      mutate(
        !!var := !!var + blur_vec,
        !!var := ifelse(abs(!!var) < mask_thres, 0, !!var)
      )
  }
  df
}

# Simple blurring function for a single value. Does the same thing as blur_it.
blur_that <- function(value, blur_abs, mask_thres) {
  as.integer(ifelse(value < mask_thres, 0, value + sample(seq(-blur_abs, blur_abs), 1)))
}

# Thanks Chuan Hong for this summary stats function!
# Sensitivity: TP/(All Positive)
# Variable names: first bit is test pos/neg, second bit is truth pos/neg
PAR.fun <- function(n11, n10, n01, n00) {
  sens <- n11 / (n11 + n01)
  spec <- n00 / (n10 + n00)
  ppv <- n11 / (n10 + n11)
  npv <- n00 / (n00 + n01)
  fscore <- 2 * sens * ppv / (sens + ppv)
  res <- c(c(sens, spec, ppv, npv, fscore, n11, n10, n01, n00))
  names(res) <- c("sens", "spec", "ppv", "npv", "fscore", "TP", "FP", "FN", "TN")
  res
}

pt_level_df <- obs_raw %>%
  mutate(selected = concept_code %in% fource_filter & 
           between(days_since_admission, -1, 1)) %>%
  group_by(patient_num) %>%
  mutate(label = if_else(max(selected) == 1, "selected", "diff")) %>%
  left_join(icu_raw, by = "patient_num") %>%
  inner_join(select(demo_raw, patient_num, admission_date), by = "patient_num")

pre_pts_byweek <- pt_level_df %>%
  select(label, patient_num, admission_date, admitted_for_covid) %>%
  distinct() %>%
  mutate(week = week(admission_date) + ((year(admission_date) - 2020) * 52)) %>%
  group_by(week, label) %>%
  summarise(
    truedot = sum(admitted_for_covid, na.rm = TRUE),
    falsedot = sum(!admitted_for_covid, na.rm = TRUE),
    n_week = n() %>% na_if(0), .groups = "drop"
  )

pts_byweek <- pre_pts_byweek %>%
  group_by(week) %>%
  summarise(n_week = sum(n_week, na.rm = TRUE), .groups = "drop") %>%
  mutate(label = "all", truedot = 0, falsedot = 0) %>%
  bind_rows(pre_pts_byweek) %>%
  mutate(
    label = label %>% fct_recode(
      "All patients" = "all",
      "Kept patients" = "selected",
      "Removed patients" = "diff"
    ),
    week_name = ymd("2020-01-01") + weeks(week)
  )

blurred_pts_byweek <- pts_byweek %>%
  # Blur truedot/falsedota by just rounding up to bins the size of the threshold
  mutate(
    truedot = round_any(truedot, threshold) + 1,
    falsedot = round_any(falsedot, threshold) + 1
  ) %>%
  blur_it("n_week", blur, threshold)

# load('4CE_utils/data/MGB_admitfilterresults.rda')
# pts_byweek = MGB_results$pts_byweek

# 8/20/21 - USE THIS TO MAKE THE FILTER IS AN **AND** FILTER
# 9/1/21 - WOULD NEED TO BE MODIFIED FOR THIS VERSION OF THE CODE
# GET RID OF THE -1, I JUST DID THIS TO ALLOW 3/4 of the list
# magic_pts_set <- obs_raw %>% filter(concept_code %in% fource_filter) %>% filter(days_since_admission<2 & days_since_admission>-2) %>% group_by(patient_num) %>% summarise(Unique_Filter_Codes=n_distinct(concept_code)) %>% filter(Unique_Filter_Codes>=length(fource_filter)-1) %>% 
#     inner_join(demo_raw,by="patient_num") %>%
#   subset(select=c(patient_num,admission_date))

```

This tells us summary stats on how many admissions are being removed and how well it performs compared to chart review. The percentage graph below shows this by week.

```{r}
# Proportion of patients with special labs vs. all pts
pct_alltime <- mean(pt_level_df$selected) * 100
print(glue("Some stats on filter: {paste(fource_filter, collapse = ',')}"))
print(glue("Percent of admissions removed by the filter: {round(pct_alltime,2)}%"))
print(summary_stats, digits = 3)
```

```{r}
# # viz
# # Reference I used for point shapes: http://www.sthda.com/english/wiki/ggplot2-point-shapes
# # And colors: http://sape.inf.usi.ch/quick-reference/ggplot2/colour
n1 <- pts_byweek %>%
  group_by(label) %>%
  summarise(sum_true = sum(truedot), sum_false = sum(falsedot))
summary_stats <- unlist(PAR.fun(n1[3, 2], n1[3, 3], n1[2, 3], n1[2, 2]))
```

```{r}
rug_df <- pts_byweek %>%
  pivot_longer(cols = c(truedot, falsedot)) %>%
  filter(!is.na(value)) %>%
  select(week_name, label, name, value) %>%
  uncount(weights = value) %>%
  mutate(name = name %>% fct_recode(
    "Correct" = "truedot",
    "Incorrect" = "falsedot"
  ))

point_df <- pts_byweek %>%
  pivot_longer(cols = c(truedot, falsedot)) %>%
  filter(!is.na(value)) %>%
  mutate(name = name %>% fct_recode(
    "Correct" = "truedot",
    "Incorrect" = "falsedot"
  ))

annotate_df <- tibble(
  label = names(summary_stats[1:2]) %>%
    recode(
      "sens" = "Sensitivity",
      "spec" = "Specificity"
    ),
  summary_stats = summary_stats[1:2],
  x = as.Date(c("2021-02-28", "2021-02-28")),
  y = c(260, 250)
) %>%
  mutate(label = glue::glue("{label}: {round(summary_stats, 2)}"))
```


```{r}
plot_trend <- function(df, cap = "") {
  df %>%
    ggplot(aes(y = n_week, x = week_name)) +
    geom_line(aes(alpha = label, linetype = label)) +
    theme_classic() +
    scale_linetype_manual(values = c(3, 2, 1)) +
    scale_alpha_discrete(range = c(0.5, 1)) +
    scale_color_brewer(palette = "Dark2", direction = -1) +
    guides(linetype = FALSE, alpha = FALSE) +
    scale_x_date(expand = expansion(c(0, 0), c(10, 90))) +
    ggtitle(paste("Patients removed by non-COVID-admission filter\n ")) +
    labs(
      color = NULL, size = "Chart reviewed",
      y = "Number of admissions by week", x = NULL,
      caption = glue("Site: {currSiteId}
                     Filter: {paste(fource_filter, collapse = ',')}
                     {cap}")
    ) +
    directlabels::geom_dl(aes(label = label),
      method = list(directlabels::dl.trans(x = x + .2), "last.bumpup")
    ) +
    theme(legend.position = "bottom") +
    geom_text(data = annotate_df, aes(x = x, y = y, label = label))
}

plot_rug <- function(df, cap = "") {
  df %>%
    plot_trend(cap) +
    geom_rug(
      data = rug_df,
      aes(x = week_name, y = NA_real_, color = name, linetype = label),
      position = "jitter"
    )
}

plot_donut <- function(df, cap = "") {
  df %>%
    plot_trend(cap) +
    geom_point(
      data = point_df %>% filter(value > 0),
      aes(y = n_week, x = week_name, size = value, color = name),
      alpha = 0.8, shape = 21, stroke = 2
    ) 
}
plot_donut(pts_byweek)
plot_rug(pts_byweek)
```
 
This shows all hospitalizations by week vs. hospitalizations by day with the filter applied ("selected"). 

Lines:
"ALL" hospitalizations includes only the hospitalizations in the COVID cohort, so all hospitalizations involve a COVID-positive test. A third line shows the difference between the two ("diff"), which is the number of hospitalizations excluded by the filter.

Dots:
The blue dots are true positives – patients who were selected by the filter and were chart reviewed as admitted for covid. 
The green dots are true negatives – patients who were not selected by the filter and were chart reviewed as not admitted for covid. 
Xs on the blue line are false positives - patients who were selected but the filter but chart reviewed as not admitted for COVID. 
Xs on the green line are false negatives - patients that were not selected by the filter but are chart reviewed as admitted for COVID.
 
```{r}
# Also plot the percentage removed by the filter
# ggplot(pct_pts, aes(y = n_week, x = week)) +
#   geom_line(color = "blue") +
#   ggtitle("Percent of patients removed by the same filter")
```

This shows the percentage of total patients seen each week that are excluded by the filter. Perhaps this is most informative, because (at MGB at least) it increases after the first wave of COVID (when hospitals were opening back up) even though the absolute number of COVID-positive hospitalizations went down.


```{r}
# Plot blurred version
plot_donut(blurred_pts_byweek, "Obfuscated")
plot_rug(blurred_pts_byweek, "Obfuscated")
```

Here we blur and threshold the results and save them to the file /4ceData/out/[siteid]_admitfilterresults.rda. Although you do not need to share anything, feel free to share the obfuscated graph or the percentage graph on slack, and/or the file, which contains only the contents of these two graphs.

```{r}
# Save Results and save the blurred version to /4ceData/out/[siteid]_admitfilterresults.rda
# TODO : We don't save the counts in the summary stats because they're not blurred. We could blur them!
results <- list(
  pts_byweek = blurred_pts_byweek,
  # pts_pct_byweek = pct_pts,
  summary_stats = summary_stats[1:5],
  pct_alltime = pct_alltime
)
site_results <- paste0(currSiteId, "_results")
assign(site_results, results)
save(list = site_results, file = file.path("/4ceData/out", paste0(currSiteId, "_admitfilterresults.rda")))
```


```{r}
# Testing playing with plotting the false positives
# pts_byweek  <- nocovid_pts_byweek_filter
# pts_byweek <- pts_byweek %>% mutate(week_name=ymd("2020-01-01")+weeks(week))
#
#  ggplot(pts_byweek,aes(y=n_week,x=week_name,group=label)) +
#   geom_line(aes(color=label,linetype=label))  +
#    ggtitle("Chart-reviewed patients removed by the hypothesized non-COVID-admission filter\n (by week)") +
#    labs(y="Number of admissions by week",x="Week (starting Jan 2020)")
```
